---
title: "NeSI-data-analysis"
author: "Fabiana"
date: "15 August 2017"
output:
  word_document: default
  html_document: default
---



```{r setupenv, echo = FALSE, eval = TRUE}

# Set up environment to read files:Need to 
#* set working directory
#* provide correct filename
#* check data has been read as a data.frame

#setwd("~/Desktop/GitHub/nesi-training/Data-analysis")
setwd("~/Documents/GitHub/nesi-training/Data-analysis")

reshed = read.csv("researchers-hed.csv") #HED institutions
rescri = read.csv("researchers-cris.csv") #CRI data
resdiscipline = read.csv("researchers-discipline.csv")
rdsurvey2016 = read.csv("rd-2016.csv")

#check it is a dataframe
#class(reshed)
#class(rescri)
#class(resdiscipline)
#class(rdsurvey2016)


library(ggplot2) #calls ggplot2 which should be installed
library(plyr) #need it for counts
library(reshape2) #need it for plotting
knitr::opts_chunk$set(fig.width=12, fig.height=8) # control figure size in output

```

## Worforce number in NZ 

Estimate number/FTE of total researchers in each Higher Ed institution, by using the fraction of total funded EP portfolios in 2012 PBRF round, and the total values provided by stats.nz 2016 R&d survey. 

According to TEC there are 9 universities and a number of other HED institutions that contribute to the PBRF 2012. All these other HED institutions are grouped into one as "all other". 

Estimate number of total tech/FTE in Higher Ed by institution by using same approach as above. 

Estimate number of total tech in CRI by distributin tech number in gov (from R&D 2016 survey) and distributing proportionately based on total number of staff in each CRI.


```{r researchers, echo=FALSE, eval = TRUE}

#read from rd-2016.csv total eft and number of researchers (exculde student researcher)
#read from rd-2016.csv total eft and number of researchers

#HED numbers

hed.researcher.number <- as.numeric(rdsurvey2016$hed.number.2016[rdsurvey2016$role == "researcher"])
hed.researcher.fte <-as.numeric(rdsurvey2016$hed.fte.2016[rdsurvey2016$role == "researcher"])
hed.tech.number<-as.numeric(rdsurvey2016$hed.number.2016[rdsurvey2016$role == "technician"])
hed.tech.fte<-as.numeric(rdsurvey2016$hed.number.2016[rdsurvey2016$role == "technician"])

#gov numbers
gov.tech.number<-as.numeric(rdsurvey2016$gov.number.2016[rdsurvey2016$role == "technician"])
gov.tech.fte<-as.numeric(rdsurvey2016$gov.fte.2016[rdsurvey2016$role == "technician"])
gov.res.number <- as.numeric(rdsurvey2016$gov.number.2016[rdsurvey2016$role == "researcher"])
gov.res.fte <- as.numeric(rdsurvey2016$gov.fte.2016[rdsurvey2016$role == "researcher"])

#split Universities from other tertiary institutions, then add other tertiary as a single line
resheduni = subset(reshed, reshed$inst.category == "heduni")
reshedother = subset(reshed, reshed$inst.category == "hedother")

reshedall<-as.data.frame(resheduni)
reshedall[nrow(resheduni) + 1,] = NA
levels(reshedall$institution) <- c(levels(reshedall$institution),"allother")
reshedall$institution[nrow(reshedall)]<-"allother"
reshedall$inst.category[nrow(reshedall)]<-"hedother"
reshedall$funded.ep[reshedall$institution=="allother"] <-sum(reshedother$funded.ep)
reshedall$student.fte[reshedall$institution=="allother"] <- sum(reshedother$student.fte)
reshedall$student.number[reshedall$institution =="allother"] <- sum(reshedother$student.number)

#Replace NA with 0
reshedother[is.na(reshedother)] <- 0
reshedall[is.na(reshedall)] <- 0
rescri[is.na(rescri)]<-0
resdiscipline[is.na(resdiscipline)] <- 0

#calculate total researchers per institution from funded eviedence porfolio numbers

total.eps <-sum(reshed$funded.ep) #total number of funded eps should be 6311.41
reshedall$percent.funded.ep <- reshedall$funded.ep * 100 /total.eps

reshedall$res.staff.number <- reshedall$percent.funded.ep * hed.researcher.number  /100
reshedall$res.staff.fte <- reshedall$percent.funded.ep * hed.researcher.fte /100

reshedall$tech.number <- reshedall$percent.funded.ep * hed.tech.number /100
reshedall$tech.fte <- reshedall$percent.funded.ep * hed.tech.fte /100

#caclulate total researchers per discipline from funded eviedence porfolio numbers
#There is no tech number per discipline

resdiscipline$percent.funded.ep <- resdiscipline$funded.ep * 100 /total.eps
resdiscipline$res.staff.number <- resdiscipline$percent.funded.ep * hed.researcher.number  /100
resdiscipline$res.staff.fte <- resdiscipline$percent.funded.ep * hed.researcher.fte /100

resdiscipline$tech.number <- as.numeric(resdiscipline$percent.funded.ep * hed.tech.number /100)
resdiscipline$tech.fte <- resdiscipline$percent.funded.ep * hed.tech.fte /100

#student numbers
# from R&D 2016
hed.student.number <- as.numeric(rdsurvey2016$hed.number.2016[rdsurvey2016$role == "researcher"])
hed.student.fte<-as.numeric(rdsurvey2016$hed.fte.2016[rdsurvey2016$role == "student.researcher"])

#From enrollments 2015
hed.student.enrolled = sum(reshedall$student.number)

#Add students, researchers and techs to total column

resdiscipline$total.number <- resdiscipline$res.staff.number + resdiscipline$student.number + resdiscipline$tech.number
resdiscipline$total.fte <- resdiscipline$res.staff.fte + resdiscipline$student.fte + resdiscipline$tech.fte

reshedall$total.number <- reshedall$res.staff.number + reshedall$student.number + reshedall$tech.number
reshedall$total.fte <- reshedall$res.staff.fte + reshedall$student.fte + reshedall$tech.fte
```


```{r learners, echo=FALSE, eval = TRUE}

#Calculate total number of NeSI learners per institution 

#add different years of NeSi learners to new totals column
reshedall$nesi.learner.total = reshedall$nesi.learner.2017 + reshedall$nesi.learner.2016 + reshedall$nesi.learner.2015 +reshedall$nesi.learner.2013

rescri$nesi.learner.total = rescri$nesi.learner.2017+rescri$nesi.learner.2016+rescri$nesi.learner.2015+rescri$nesi.learner.2013

#Calculate the proportion of number of Nesi learners per institution
#proportion =  nesi total *100/ total workforce (student, researchers and tech)
reshedall$nesi.learner.proportion = reshedall$nesi.learner.total * 100 / reshedall$total.number
rescri$nesi.learner.proportion = rescri$nesi.learner.total *100 / rescri$total.number

```

## Distribution of NeSi learners across institutions (total and per year)

```{r plot distributions, echo=FALSE, eval = TRUE}

cat("Distribution of NeSi learners by HED institutions (all years) ")

ggplot(reshedall, aes(x=reorder(institution,nesi.learner.total), y=nesi.learner.total)) +
  geom_bar(stat="identity") + 
  #ggtitle("# NeSI Learners by institution (HED)")+
  ylab("Total number of learners")+
  xlab("")+
  coord_flip()

cat("Distribution of NeSI learners by HED institutions (broken by years)")

hedperyear <- melt(reshedall[,c('institution','nesi.learner.2013','nesi.learner.2015','nesi.learner.2016','nesi.learner.2017')],id.vars = 1)
ggplot(hedperyear,aes(x = institution,y = value)) + 
  geom_bar(aes(fill = variable),stat = "identity",position = "dodge")  +
  ylab("number of learners") +
  xlab("")+
  coord_flip()

cat("Distribution of NeSI learners by CRIs (all years)")

ggplot(rescri, aes(x=reorder(institution,nesi.learner.total), y=nesi.learner.total)) +
  geom_bar(stat="identity") + 
  ylab("Total number of learners")+
  xlab("")+
  coord_flip()
  
cat("Distribution of NeSI learners by CRI (broken by years)")

criperyear <- melt(rescri[,c('institution','nesi.learner.2013','nesi.learner.2015','nesi.learner.2016','nesi.learner.2017')],id.vars = 1)
ggplot(criperyear,aes(x = institution,y = value)) + 
  geom_bar(aes(fill = variable),stat = "identity",position = "dodge")  +
  ylab("number of learners") +
  xlab("")+
  coord_flip()

```

## NeSI learners as a proportion of workforce

```{r nesi learner proportion, echo=FALSE, eval=TRUE}
cat("Proportion of NeSi learners by HED institutions (all years) ")

ggplot(reshedall, aes(x=reorder(institution,nesi.learner.proportion), y=nesi.learner.proportion)) +
  geom_bar(stat="identity") + 
  #ggtitle("# NeSI Learners % by institution (HED)")+
  ylab("% of workforce")+
  xlab("")+
  coord_flip()

cat("Distribution of % NeSI learners by CRIs (all years)")

rescrionly = subset(rescri, rescri$inst.category == "cri")
ggplot(rescrionly, aes(x=reorder(institution,nesi.learner.proportion), y=nesi.learner.proportion)) +
  geom_bar(stat="identity") + 
  ylab("Total number of learners")+
  xlab("")+
  coord_flip()

```


##Proportion of the workforce in different disciplines that use NeSI

```{r nesi user proportion, echo=FALSE, eval=TRUE}

#calculate percent of nesi users per total estimated workforce
# number nesi users *100 / total workforce

#NeSI users by discipline

discipline.nesi.users = sum(resdiscipline$nesi.users) 
discipline.res.number = sum(resdiscipline$res.staff.number) + sum(resdiscipline$student.number) + sum(resdiscipline$tech.number)


#total research workforce (per rd 2016) - No info about discipline
#Need to recalculate this .....

res.gov.2016 = gov.res.number + gov.tech.number
res.hed.2016 = hed.researcher.number + hed.student.number + hed.tech.number
hed.gov.2016 = res.gov.2016 + res.hed.2016 #HED + Government total users

percent.nesi.users.gov.hed = discipline.nesi.users * 100 / hed.gov.2016
percentage.nesi.users = discipline.nesi.users *100 /discipline.res.number
resdiscipline$nesi.percentage = resdiscipline$nesi.users * 100 / resdiscipline$total.number

cat("There are", hed.gov.2016, "researchers in higer education and government combined, and ", discipline.nesi.users, "total nesi users. That is, NeSi users are about ", format(percent.nesi.users.gov.hed, digits = 3, nsmall =2),"% of the gov + HED workforce. The workforce for business has been excluded from the calculation. This includes support staff - so perhaps I should remove this.") 

cat("A calculation of the percentage of the workforce by discipline cannot be done with the avilable data - but it can be calculated assuming all nesi users are in HED - this provides an overestimate of the percentage of workforce within the discipline that is using NeSI. ")

ggplot(resdiscipline, aes(x=reorder(discipline,nesi.percentage), y=nesi.percentage)) +
  geom_bar(stat="identity") + 
  xlab("") +
  ylab("users as percentage of workforce")+
  ggtitle("% of workforce using NeSI, by discipline")+
  coord_flip()

```

This is, however, an overestimation of the actual proportions since it assumes all NeSI users come from the HED sector. I am unable to find a breakdown of research force within CRIs that I can map onto disciplines. The only way around this would be to assume that the distribution across disciplines in the government and higher education are the same and break up the government data across those proportions - but this would probably be a really incorrect assumption. 



```{r donut plot, echo=FALSE, eval=FALSE}

# Create data to plot
dat <- data.frame(institution = resinstitution$institution, staffnumber = resinstitution$res.staff.number, insttype = resinstitution$inst.category)

# Add addition columns, needed for drawing with geom_rect.
dat$fraction = dat$staffnumber[dat$insttype=="heduni"] / sum(dat$staffnumber[dat$insttype=="heduni"])
dat = dat[order(dat$fraction[dat$insttype=="heduni"]), ]
dat$ymax = cumsum(dat$fraction[dat$insttype=="heduni"])
dat$ymin = c(0, head(dat$ymax[dat$insttype=="heduni"], n=-1))
 
# Make the plot
p2 = ggplot(dat, aes(fill=institution, ymax=ymax, ymin=ymin, xmax=4, xmin=3)) +
     geom_rect(colour="grey30") +
     coord_polar(theta="y") +
     xlim(c(0, 4)) +
     theme_bw() +
     theme(panel.grid=element_blank()) +
     theme(axis.text=element_blank()) +
     theme(axis.ticks=element_blank()) +
     ggtitle("proportion of research staff at universities") 
p2
```

